The Astropy Project

### 2017-11-20 main.tex
Improve text

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
% heavily edited by Adrian PW - let me know if you want to revert anything!
All astronomical research makes use of software in some way.  Astronomy as a field has thus long supported the development of software tools
for astronomical tasks: from scripts that enable individual scientific research to software packages for small collaborations to data reduction pipelines for survey operations.
Some software packages are or were supported by large institutions and are intended for a wide range of users. These packages typically provide some level of documentation and user support or
training.
Other packages are developed by individual researchers or research groups and
are then typically used by smaller groups for more domain-specific purposes.
Whether for a package meant for wide distribution or for scripts and programs
for a specific research project, the implementation of astronomical software can
be eased through the use of a library that provides core functionality that is
common to many astronomical tasks.
The users of such software then also benefit from a community and ecosystem
built around a common foundation.
The \astropy project has grown to become this community for \python astronomy
software, and the \astropypkg core package a feature-rich \python library.

%\inlinecomment{Tom R}{the 'this community' above is a bit too general - do we mean it is the only community in astronomy, or in the Python astronomy world? I'd change this to either 'to become this community for Python astronomy software' or 'to become such a community'. For instance yt had a community before we came along for the simulation community, so I want to make sure we don't claim we are the only community out there.}
%\inlinecomment{Perry G}{Agreed}

The development of the \astropypkg core package began as a largely
community-driven effort to standardize core functionality for astronomical
software in \python.
In this way, its genesis differs from but builds upon many substantial and
% * <bsipocz@gmail.com> 2017-11-13T13:13:08.884Z:
%
% how do astroy builds upon these? or it's more like builds upon the developer experience?
%
%
% ^.
former astronomical software development efforts that were commissioned or
initiated through large institutional support, for example IRAF \citep[developed
at NOAO;][]{IRAF}, MIDAS \citep[developed at ESO;][]{MIDAS}, or Starlink
\citep[originally developed by a consortium of UK institutions and now
maintained by the East Asian Observatory;][]{starlink1982,starlink2013}.
More recently, community-driven efforts have seen significant success in the astronomical sciences \citep{yt}
%\inlinecomment{BMS}{I like the idea to reference yt here}
%
\python\footnote{\url{https://www.python.org/}} is an increasingly popular,
general-purpose programming language that is available under a permissive open
source software license free of charge for all major operating systems.
Stable and well-developed packages provide support for array representation and arithmetic
\citep[\package{numpy};][]{numpy}, a wide variety of functions for scientific
computing \citep[\package{scipy};][]{numpy}, and publication-quality plotting
\citep[\package{matplotlib};][]{matplotlib}. Tens of thousands of other packages
are available which can help with tasks that are not astronomy specific but
might be performed in the course of astronomical research, for example interfacing with databases or statistical inference.

@@

As of version 2.0, \astropypkg contains $212244$ lines of code\footnote{This
line count includes comments, as these are often as important for
maintainability as the code itself.  Without comments there are $142197$ lines
of code.} contributed by $232$ unique contributors over $19270$ \texttt{git}
commits.
The commit distribution \inlinecomment{Tom R}{too vague, maybe 'the distribution of number of commits as a function of ...'} follows an approximate power-law (with log-slope of
$\approx -0.47$) \inlinecomment{AG}{round to -0.5?  This is astronomy, after all.}, meaning most of the commits were and are contributed by a
relatively small fraction of the contributor pool.
Roughly half of the commits were contributed by the top $4$
contributors, while $\sim 90$ percent of the commits came from the top $24$
contributors.

@@

Central to the success of \astropy is an open environment where anybody can
contribute to the project.
However, this model leads to ``organic'' growth where different features are
implemented by different people with different programming styles and
interfaces.
Thus, \astropy has a mechanism to more formally propose significant changes to
the core package (e.g., re-writing the coordinates subpackage; \citealt{ape5}),
to plan out major new features (e.g., a new file format; \citealt{ape6}), or
institute new organization-wide policies (e.g., adopting a code of conduct;
\citealt{ape8}).
This mechanism is called ``Astropy Proposal for Enhancement'' (APE) and are
modeled after the ``Python Enhancement Proposals'' (PEP) that guide the
development of the \python programming language.
In an APE, one or more authors describe in detail the proposed changes or
additions, including a rationale for the changes, how these changes will be
implemented, and, in the case of code, what the interface will be \citep{ape1}.
The APEs are discussed and refined by the community before much work is invested
into a detailed implementation; anyone is welcome to contribute to these
discussions during the open consideration period. APEs are proposed via pull
requests on a dedicated GitHub repository\footnote{\url{https://github.com/astropy/astropy-APEs}},
and anyone can therefore read the proposed APEs and leave comments in-line.
In previous APEs a community consensus emerged and APEs are accepted and become
the basis for future work at this point.
In cases where consensus cannot be reached, the
\astropy coordination committee can decide to close the discussion and
make an executive decision based on the community input on the APE in question.

@@

@@

The \astropypkg core package is available from several of the \python distributions for scientific
% * <mattwcraig@gmail.com> 2017-10-27T07:13:51.093Z:
%
% > \astropypkg i
% I changed a few of the macros in here to refer to the package rather than the project because it is the package being discussed here.
%
% ^.
computing, such as Anaconda or Enthought Canopy, which include also some of
\astropy's affiliated packages.
Alternatively, \astropypkg can be readily installed from the Python Package
Index\footnote{\url{https://pypi.python.org/pypi}} (PyPI)
\inlinecomment{Tom R}{make footnote link directly to astropy on PyPI?}
using \texttt{pip}, the standard command-line tool for installing \python packages.
For most recent \astropypkg releases, pre-built wheel distributions are obtainable
from PyPI for Windows, Mac OS X and Linux systems that provide faster
installation compared to the source distribution.
% hamogu: I think that's too detailed for a journal paper.
%     It's in the docs if anyone needs it.
%Optionally, the version of
%the package and the dependencies needed to run or build the software are given
%by a specifier to the pip package manager as described in PEP 440\footnote{For
%more detailed installation instructions and optional flags see
%\url{http://docs.astropy.org/en/stable/install.html\#installing-astropy}}.

@@

\subsection{Support of Astropy}

The \astropy project, as of the v2.0 release, does not receive any direct financial support for
the development of the \astropypkg.  Development of the \astropypkg is supported
through contributions by individuals to work on astropy in their own personal
time, contributions from individuals or groups to work on \astropy as 
part of a research project, or contributions from institutes that allocate 
people to work on \astropy. 
A list of organizations that have contributed to \astropy
are listed in the Acknowledgements.  
Different models have been proposed for the support of astropy 
(ie. \cite{2016arXiv161003159M}), 
but a sustainable, community-based model has not yet been established.
The astropy project has the ability to accept financial contributions 
(individual or institutes) through the NumFOCUS organization\footnote{NumFOCUS 
is a 501(c)(3) nonprofit that supports and promotes world-class, innovative, 
open source scientific computing.},  
which generally covers the direct costs incurred by the \astropy project
along with supporting

% Removing unless this gets written about in
% a later version
% This section should be a straight forward/factual description of the current
% funding and support provided to astropy with the purpose of explaining the
% current structure to the community

@@

@@

The benchmarks are meant to evolve over time and include an increasing variety of cases. At the moment, the benchmarks are set up as follows - we have generated a standard set of 1000 pairs of random longitudes/latitudes that we use in all benchmarks. Each benchmark is then defined using an input and output coordinate frame, using all combinations of FK4, FK5, Galactic, ICRS and Ecliptic frames. For now we set the epoch of observation to J2000, and the frame to J2000 in the case of the FK5 and ecliptic frames and B1950 for the FK4 frame, but in future we plan to include a larger variety of epochs and equinoxes, as well as tests of conversion to/from Altitude/Azimuth. For each benchmark, we convert the 1000 longitudes/latitudes from the input/output frame with all tools and quantify the comparison by looking at the median, minimum, maximum, and standard deviation of the absolute separation of the output coordinates from each pair of tools. \tablename~\ref{tab:coordinate_benchmarks} gives an example of the relative accuracy of the conversion from FK4 to Galactic coordinates for all pairs of tools. This shows for example that \astropy, Kapteyn and PyTPM agree perfectly (to the precision shown), while PALpy, pySLALIB, and PyAST also agree perfectly amongst themselves, but show an offset of around 0.2\arcsec with the former three packages. Finally, PyEphem disagrees with other packages by 0.4--0.8\arcsec. These values are only meant to be illustrative and will change over time as the benchmarks are refined and packages are updated.
%\inlinecomment{Perry G}{Wouldn't the table be better represented as a matrices? Or perhaps as a matrix as a crude image with colors or intensities representing the size of the quantity. Is the difference between median and mean significant enough to warrant showing both? A link to the actual numbers may satisfy readers that want the exact values, but I suspect the majority don't need to see that precision.}
\begin{table}
\begin{center}
\begin{tabular}{llcccc}
\hline
\hline
Package 1 & Package 2 & Median & Mean & Maximum & Std. Dev. \\
          &           & arcsec & arcsec & arcsec & arcsec \\
\hline
astropy  &  kapteyn  & 0.000  & 0.000 & 0.000 & 0.000 \\
\nodata  &  palpy    & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pyast    & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pyephem  & 0.459  & 0.458 & 0.860 & 0.231 \\
\nodata  &  pyslalib & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pytpm    & 0.000  & 0.000 & 0.000 & 0.000 \\
kapteyn  &  palpy    & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pyast    & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pyephem  & 0.459  & 0.458 & 0.860 & 0.231 \\
\nodata  &  pyslalib & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pytpm    & 0.000  & 0.000 & 0.000 & 0.000 \\
palpy    &  pyast    & 0.000  & 0.000 & 0.000 & 0.000 \\
\nodata    &  pyephem  & 0.563  & 0.570 & 1.012 & 0.253 \\
\nodata    &  pyslalib & 0.000  & 0.000 & 0.000 & 0.000 \\
\nodata    &  pytpm    & 0.218  & 0.196 & 0.249 & 0.056 \\
pyast    &  pyephem  & 0.563  & 0.570 & 1.012 & 0.253 \\
\nodata    &  pyslalib & 0.000  & 0.000 & 0.000 & 0.000 \\
\nodata    &  pytpm    & 0.218  & 0.196 & 0.249 & 0.056 \\
pyephem  &  pyslalib & 0.563  & 0.570 & 1.012 & 0.253 \\
\nodata  &  pytpm    & 0.459  & 0.458 & 0.860 & 0.231 \\
pyslalib &  pytpm    & 0.218  & 0.196 & 0.249 & 0.056 \\
\hline
\end{tabular}
\end{center}
\caption{Comparison of the accuracy of the FK4 to Galactic transformation between different packages.\label{tab:coordinate_benchmarks}}
\end{table}

@@

\subsubsection{Barycentric and Heliocentric corrections}
Detailed eclipse or transit
        timing requires accounting for light travel time differences from the
        source to the observatory because of the Earth's motion.
        It is therefore common to instead convert times to the Solar System
        barycenter or heliocenter where the relative timing of photons is
        standardized.
        With the location of a source on the sky (i.e. a \texttt{SkyCoord}
        object), the location of an observatory on Earth (i.e. an
        \texttt{EarthLocation} object), and time values as \texttt{Time}
        objects, the time corrections to shift to the solar system barycenter or
        heliocenter can now be computed with \package{astropy.time} using the
        \texttt{light\_travel\_time} method of a \texttt{Time} object.

@@

\inlinecomment{Perry G}{Some mention that some of these are inspired by Pandas?}

\subsubsection{Support for grouped table operations}

A table can contain data that naturally forms groups -- for example it may
contain multiple observations of a few sources at different points in time
and in different bands. We may then want to split the table into groups based
on the combination of source observed and the band, and then combine the
results for each combination of source and band in some way (for example finding
the mean or standard deviation of the fluxes or magnitudes over time) or filter
the groups based on user-defined criteria. These kinds of grouping and
aggregation operations are now fully supported by \texttt{Table} objects.

\subsubsection{Support for table concatenation}

\texttt{Table} objects can now be combined in several different ways. If two
tables have the same rows, we may want to stack them ``vertically'' to create a
new table with the same columns but all rows. If two tables are row-matched but
have distinct columns, we may want to stack them ``horizontally'' to create a
new table with the same rows but all columns. For other situations, more general
table concatenations or joins are also possible when two tables share some
common columns.

\subsubsection{Using astropy objects in tables}

The \texttt{Table} object now allows \texttt{Quantity} objects, celestial
coordinate objects (\texttt{SkyCoord}), and date/time objects (\texttt{Time}) to
be used as columns, and also provides a general way for other user-defined
objects to be used as columns. This makes it possible for example to easily
represent catalogs of sources or time series as in Astropy, and having both the
benefits of the \texttt{Table} object (such as accessing specific rows/columns
or groups of rows/columns, combining tables, and so on) and the benefits of e.g.
the \texttt{SkyCoord} or \texttt{Time} classes (such as converting the
coordinates to a different frame, or accessing the date/time as a modified
Julian date or in another time scale).

@@

\label{sec:infrastructure}
%\inlinecomment{BMS}{should we describe astropy-helpers even though it's future is unsure?}
% Text also edited by Adrian PW

@@

%\inlinecomment{BMS}{should we mention the refactored infra here?  e.g. pytest-astropy and sphinx-astropy?}

@@

Beyond the core code, the \astropy project is also further developing the
\astropy-managed affiliated packages.
While these may not be integrated into the \astropypkg core package, these
projects provide code that is useful to the astronomical community and meet the
testing and documentation standards of \astropy.
Some of these new efforts includes an initiative to develop tools for
spectroscopy \citep[\package{specutils}, \package{specreduc}, \package{specviz}]
{ape13}, integration of LSST software, and packages to support HEALPIX
projection. %\citep{astropy-healpix}
%\inlinecomment{BMS}{TODO: uncomment references of packages once they are available}

@@

%\inlinecomment{SMC}{Can a brief description of each code be added?}

>>>>

\section{Introduction} \label{sec:intro}
% first draft by Moritz (hamogu)
% heavily edited by Adrian PW - let me know if you want to revert anything!
All astronomical research makes use of software in some way.  Astronomy as a field has thus long supported the development of software tools
for astronomical tasks: from scripts that enable individual scientific research to software packages for small collaborations to data reduction pipelines for survey operations.
Some software packages are or were supported by large institutions and are intended for a wide range of users. These packages typically provide some level of documentation and user support or
training.
Other packages are developed by individual researchers or research groups and
are then typically used by smaller groups for more domain-specific purposes.
Whether for a package meant for wide distribution or for scripts and programs
for a specific research project, the implementation of astronomical software can
be eased through the use of a library that provides core functionality that is
common to many astronomical tasks.
The users of such software then also benefit from a community and ecosystem
built around a common foundation.
The \astropy project has grown to become this community, and the \astropypkg
core package a feature-rich library for \python-based astronomical software.

\inlinecomment{Tom R}{the 'this community' above is a bit too general - do we mean it is the only community in astronomy, or in the Python astronomy world? I'd change this to either 'to become this community for Python astronomy software' or 'to become such a community'. For instance yt had a community before we came along for the simulation community, so I want to make sure we don't claim we are the only community out there.}
\inlinecomment{Perry G}{Agreed}

The development of the \astropypkg core package began as a largely
community-driven effort to standardize core functionality for astronomical
software in \python.
In this way, its genesis differs from but builds upon many substantial and
% * <bsipocz@gmail.com> 2017-11-13T13:13:08.884Z:
%
% how do astroy builds upon these? or it's more like builds upon the developer experience?
%
%
% ^.
former astronomical software development efforts that were commissioned or
initiated through large institutional support, for example IRAF \citep[developed
at NOAO;][]{IRAF}, MIDAS \citep[developed at ESO;][]{MIDAS}, or Starlink
\citep[originally developed by a consortium of UK institutions and now
maintained by the East Asian Observatory;][]{starlink1982,starlink2013}.
% More recently, community-driven efforts have seen significant success in the astronomical sciences \citep{yt}
\inlinecomment{BMS}{I like the idea to reference yt here}
%
\inlinecomment{Tom R}{Python calls itself a programming language or general-purpose language, not specifically scripting language. I would remove 'scripting' below.}
\inlinecomment{Perry G}{Yes, it's time to drop the scripting adjective; it's far more general than that now.}
\python\footnote{\url{https://www.python.org/}} is an increasingly popular,
general-purpose scripting language that is available under a permissive open
source software license free of charge for all major operating systems.
Stable and well-developed packages provide support for array representation and arithmetic
\citep[\package{numpy};][]{numpy}, a wide variety of functions for scientific
computing \citep[\package{scipy};][]{numpy}, and publication-quality plotting
\citep[\package{matplotlib};][]{matplotlib}. Tens of thousands of other packages
are available which can help with tasks that are not astronomy specific but
might be performed in the course of astronomical research, for example interfacing with databases or statistical inference.

@@

As of version 2.0, \astropypkg contains $212244$ lines of code\footnote{This
line count includes comments, as these are often as important for
maintainability as the code itself.  Without comments there are $142197$ lines
of code.} contributed by $232$ unique contributors over $19270$ \texttt{git}
commits. We can divide up the commits by
contributor and construct a distribution which is the number of commits
for each contributor versus the rank of the contributor (where 1 is the
contributor who has made the most commits, 2 is the contributor with the
second most commits, and so on). When doing this, we find that this distribution
follows an approximate power-law (with log-slope of
$\approx -0.5$), meaning most of the commits were and are contributed by a
relatively small fraction of the contributor pool.
Roughly half of the commits were contributed by the top $4$
contributors, while $\sim 90$ percent of the commits came from the top $24$
contributors.

@@

Central to the success of \astropy is an open environment where anybody can
contribute to the project.
However, this model leads to ``organic'' growth where different features are
implemented by different people with different programming styles and
interfaces.
Thus, \astropy has a mechanism to more formally propose significant changes to
the core package (e.g., the release cycle; \citealt{ape2})\inlinecomment{AG}{Is the release cycle really part of the core package?  That's how this reads.  Is there something more concretely linked to the code we can use as an example here?}, to plan out major
new features, or institute new organization-wide policies (e.g.,
code of conduct; \citealt{ape8}).
This mechanism is called ``Astropy Proposal for Enhancement'' (APE) and are
modeled after the ``Python Enhancement Proposals'' (PEP) that guide the
development of the \python programming language.
In an APE, one or more authors describe in detail the proposed changes or
additions, including a rationale for the changes, how these changes will be
implemented, and, in the case of code, what the interface will be \citep{ape1}.
The APEs are discussed and refined by the community before much work is invested
into a detailed implementation; anyone is welcome to contribute to these
discussions during the open consideration period by leaving comments on the
respective pull request.\footnote{\url{https://github.com/astropy/astropy-APEs}} \inlinecomment{Tom R}{first mention of a pull request in the context of an APE - need to make it clear that APEs are proposed by PRs}
In previous APEs a community consensus emerged and APEs are accepted and become
the basis for future work at this point.
In cases where consensus cannot be reached, the
\astropy coordination committee can decide to close the discussion and
make an executive decision based on the community input on the APE in question.

@@

%\subsection{Mixin}

@@

The \astropypkg core package is available from several of the \python distributions for scientific
% * <mattwcraig@gmail.com> 2017-10-27T07:13:51.093Z:
%
% > \astropypkg i
% I changed a few of the macros in here to refer to the package rather than the project because it is the package being discussed here.
%
% ^.
computing, such as Anaconda or Enthought Canopy, which include also some of
\astropy's affiliated packages.
Alternatively, \astropypkg can be readily installed from the Python Package
Index\footnote{\url{https://pypi.python.org/pypi}} (PyPI) 
\inlinecomment{Tom R}{make footnote link directly to astropy on PyPI?}
using pip, the
recommended command-line tool for installing \python packages
that is included by default since \python 2 $\geq$ 2.7.9 and \python 3 $\geq$ 3.4.

\inlinecomment{Tom R}{is this actually true? I installed Python 3.6 from source recently and it didn't have the pip command readily available. What it does have is the ensurepip module, but we might want to be careful how we phrase this here. Maybe just remove mention of pip being included by default?}
For most recent \astropypkg releases, pre-built wheel distributions are obtainable
from PyPI for Windows, Mac OS X and Linux systems that provide faster
installation compared to the source distribution.
% hamogu: I think that's too detailed for a journal paper.
%     It's in the docs if anyone needs it.
%Optionally, the version of
%the package and the dependencies needed to run or build the software are given
%by a specifier to the pip package manager as described in PEP 440\footnote{For
%more detailed installation instructions and optional flags see
%\url{http://docs.astropy.org/en/stable/install.html\#installing-astropy}}.

@@

%\subsection{Support of Astropy}
% Removing unless this gets written about in
% a later version
% This section should be a straight forward/factual description of the current
% funding and support provided to astropy with the purpose of explaining the
% current structure to the community

@@

\subsubsection{Using quantities in tables}

\inlinecomment{Tom R}{this is a new feature of table, not units, so I think we don't want to mention this here}

        The \texttt{astropy.table} subpackage (see
        \sectionname~\ref{sec:table}) now supports using quantities in tables
        with the \texttt{QTable} class.
        As a result, units specified in FITS tables are supported, and tables
        with quantities can be read and written to various file formats.

@@

The benchmarks are meant to evolve over time and include an increasing variety of cases. At the moment, the benchmarks are set up as follows - we have generated a standard set of 1000 pairs of random longitudes/latitudes that we use in all benchmarks. Each benchmark is then defined using an input and output coordinate frame, using all combinations of FK4, FK5, Galactic, ICRS and Ecliptic frames. For now we set the epoch of observation to J2000, and the frame to J2000 in the case of the FK5 and ecliptic frames and B1950 for the FK4 frame, but in future we plan to include a larger variety of epochs and equinoxes, as well as tests of conversion to/from Altitude/Azimuth. For each benchmark, we convert the 1000 longitudes/latitudes from the input/output frame with all tools and quantify the comparison by looking at the median, minimum, maximum, and standard deviation of the absolute separation of the output coordinates from each pair of tools. \tablename~\ref{tab:coordinate_benchmarks} gives an example of the relative accuracy of the conversion from FK4 to Galactic coordinates for all pairs of tools. This shows for example that \astropy, Kapteyn and PyTPM agree perfectly (to the precision shown), while PALpy, pySLALIB, and PyAST also agree perfectly amongst themselves, but show an offset of around 0.2\arcsec with the former three packages. Finally, PyEphem disagrees with other packages by 0.4--0.8\arcsec. These values are only meant to be illustrative and will change over time as the benchmarks are refined and packages are updated.
\inlinecomment{Perry G}{Wouldn't the table be better represented as a matrices? Or perhaps as a matrix as a crude image with colors or intensities representing the size of the quantity. Is the difference between median and mean significant enough to warrant showing both? A link to the actual numbers may satisfy readers that want the exact values, but I suspect the majority don't need to see that precision.}
\begin{table}
\begin{center}
\begin{tabular}{llcccc}
\hline
\hline
Package 1 & Package 2 & Median & Mean & Maximum & Std. Dev. \\
          &           & arcsec & arcsec & arcsec & arcsec \\
\hline
astropy  &  kapteyn  & 0.000  & 0.000 & 0.000 & 0.000 \\
\nodata  &  palpy    & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pyast    & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pyephem  & 0.459  & 0.458 & 0.860 & 0.231 \\
\nodata  &  pyslalib & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pytpm    & 0.000  & 0.000 & 0.000 & 0.000 \\
kapteyn  &  palpy    & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pyast    & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pyephem  & 0.459  & 0.458 & 0.860 & 0.231 \\
\nodata  &  pyslalib & 0.218  & 0.196 & 0.249 & 0.056 \\
\nodata  &  pytpm    & 0.000  & 0.000 & 0.000 & 0.000 \\
palpy    &  pyast    & 0.000  & 0.000 & 0.000 & 0.000 \\
\nodata    &  pyephem  & 0.563  & 0.570 & 1.012 & 0.253 \\
\nodata    &  pyslalib & 0.000  & 0.000 & 0.000 & 0.000 \\
\nodata    &  pytpm    & 0.218  & 0.196 & 0.249 & 0.056 \\
pyast    &  pyephem  & 0.563  & 0.570 & 1.012 & 0.253 \\
\nodata    &  pyslalib & 0.000  & 0.000 & 0.000 & 0.000 \\
\nodata    &  pytpm    & 0.218  & 0.196 & 0.249 & 0.056 \\
pyephem  &  pyslalib & 0.563  & 0.570 & 1.012 & 0.253 \\
\nodata  &  pytpm    & 0.459  & 0.458 & 0.860 & 0.231 \\
pyslalib &  pytpm    & 0.218  & 0.196 & 0.249 & 0.056 \\
\hline
\end{tabular}
\end{center}
\caption{Comparison of the accuracy of the FK4 to Galactic transformation between different packages.\label{tab:coordinate_benchmarks}}
\end{table}

@@

\subsubsection{Barycentric and Heliocentric corrections}
Detailed eclipse or transit
        timing requires accounting for light travel time differences from the
        source to the observatory because of the Earth's motion.
        It is therefore common to instead convert times to the Solar System
        barycenter or heliocenter where the relative timing of photons is
        standardized.
        With the location of a source on the sky (i.e. a \texttt{SkyCoord}
        object), the location of an observatory on Earth (i.e. an
        \texttt{EarthLocation} object), and time values as \texttt{Time}
        objects, the time corrections to shift to the solar system barycenter or
        heliocenter can now be computed with \package{astropy.time} using the
        \texttt{light\_travel\_time} method of a \texttt{Time} object.
\subsubsection{\texttt{Time} objects as \texttt{Table} columns} \texttt{Time} objects
        can now be added as columns in \package{astropy.table} \texttt{Table}
        objects. This is described in more detail in
        \sectionname~\ref{sec:table} with the concept of a ``mixin'' column
        type.

\inlinecomment{Tom R}{I would remove this tiny section. As for units, I think we should only mention the whole mixin stuff in the table section, not for every single object that can go in a table}

@@

\inlinecomment{Perry G}{Some mention that some of these are inspired by Pandas?}
\subsubsection{Support for grouped table operations}
Tables can contain data that naturally form groups, for example monitoring of an object in different bands.
We may then want to split the table or a column into these groups (in this example: observations in different filters)  and
compute some statistic on the data, or filter the groups based on some
criteria.
Grouped operations are now supported by \texttt{Table} objects.
%,
%mainly through the \texttt{group\_by()} method and subsequent operations
%on the grouped version of the table or column.

\subsubsection{Support for table concatenation}
Tables can be combined in several
different ways.
If two tables have the same rows, we may want to stack them
``vertically'' to create a new table with the same columns but all rows.
If two tables are row-matched but have distinct columns, we may want to
stack them ``horizontally'' to create a new table with the same rows
but all columns.
For other situations, more general table concatenations or joins are
also possible when two tables share some common columns.
%All three of these operations are now possible with \texttt{Table}
%objects through the \texttt{vstack}, \texttt{hstack}, and \texttt{join}
%functions in \package{astropy.table}.

\subsubsection{Mixin columns for time and coordinates}
The \texttt{Table} object now supports \texttt{Quantity}, sky coordinate (\texttt{SkyCoord}),
astronomical \texttt{Time} objects and and other objects to serve
as array-like containers for column data through the ``mixin'' protocol.

\inlinecomment{Tom R}{this is waaaaay too short and jargony. It's a very cool feature that deserves more reader-friendly text!}

%hamogu: I've commented this out.
% I think it's very specific and not relevant to non-developers.
% If we want to keep it, we need to add at least 102 senteces about what Pandas
% is and I think that's more text than is warrented for this feature.

%\subsubsection{Transform to and from \texttt{Pandas} \texttt{DataFrame} objects}
%\texttt{Table} objects can now be created from a \texttt{DataFrame}
%object (\texttt{from\_pandas()}), or exported to a \texttt{DataFrame}
%object (\texttt{to\_pandas()}).
%\texttt{DataFrame} objects do not support multi-dimensional array
%columns, so this conversion will fail if the source \texttt{Table}
%object contains any such columns.

@@

\label{sec:infrastructure}
\inlinecomment{BMS}{should we describe astropy-helpers even though it's future is unsure?}
% Text also edited by Adrian PW

@@

\inlinecomment{BMS}{should we mention the refactored infra here?
  e.g. pytest-astropy and sphinx-astropy?}

@@

Beyond the core code, the \astropy project is also further developing the
\astropy-managed affiliated packages.
While these may not be integrated into the \astropypkg core package, these
projects provide code that is useful to the astronomical community and meet the
testing and documentation standards of \astropy.
Some of these new efforts includes an initiative to develop tools for
spectroscopy \citep[\package{specutils}, \package{specreduc}, \package{specviz}]
{ape13}, integration of LSST software, and packages to support HEALPIX
projection. %\citep{astropy-healpix}
\inlinecomment{BMS}{TODO: uncomment references of packages once they are available}

@@

\inlinecomment{SMC}{Can a brief description of each code be added?}
